Kindly Bent to Free Us

### 2020-05-26 kindly/motivation.tex
folding

% We have seen that borrows are limited to a region.
In the examples so far, regions do not appear in type signatures.
But for certain programming idioms, we want to extend the scope of a
region across a function boundary.
For instance, how should we fold on an array of linear objects?
A naive fold function would only work on unrestricted content:
\begin{lstlisting}
val fold : ('a : un) => ('a -> 'b -> 'b) -> 'a Array.t -> 'b -> 'b
\end{lstlisting}

If we want to work on linear and affine resources,
we must make sure those resources are not leaked in the result.
content. We obtain the following signature:
\begin{lstlisting}
val fold :
  ('b:'k),('k <= aff_r) =>
  (&(aff_r1,'a) -> 'b -{aff_r1}> 'b) -> &('k_1,'a Array.t) -> 'b -{'k_1}> 'b
\end{lstlisting}

@@

As an easy example, we fold over an array of files to compute the sum of their sizes:
\begin{lstlisting}
let total_size = fold (fun f s -> File.size f + s) &all_files 0
\end{lstlisting}

This approach is not sufficient if we want to mutate the array during iteration.
To this end, we need to take an exclusive borrow of the structure to iterate
on:
\begin{lstlisting}
val iter_mut : &!('a Array.t) -> (&!(aff_r1,'a) -{un}> unit) -{aff}> unit
\end{lstlisting}

>>>>

% We have seen that borrows are limited to a region.
In the examples so far, regions do not appear in type signatures.
But for certain programming idioms, we want to extend the scope of a
region across a function boundary.
For instance, how should we fold on an array of linear objects?
Here is a first attempt at the type of a fold function:
\begin{lstlisting}
val fold : ('a : 'k) => ('a -> 'b -{'k}> 'b) -> 'a Array.t -> 'b -{lin}> 'b
\end{lstlisting}
This type puts no restrictions on the element type of the array, but
it requires the @fold@ function to consume the array and all its
elements. The last function arrow is linear because the array type (from
\cref{sec:imper-progr}) is linear.

If we want to work on borrows of linear and affine resources, then the
typing gets more involved because
we must make sure those borrows are not leaked in the result.
We obtain the following signature for @bfold@, the borrowing fold operation:
\begin{lstlisting}
val bfold : ('b:'k),('k <= aff_r) => (&(aff_r1,'a) -> 'b -{aff_r1}> 'b) -> &('k_1,'a Array.t) -> 'b -{'k_1}> 'b
\end{lstlisting}

@@

As an easy example, we fold over an array of files
@all_files : File.t Array.t@ to compute the sum of their sizes:
\begin{lstlisting}
let total_size_of files = bfold (fun f s -> File.size f + s) files 0
let total_size = total_size_of &all_files
\end{lstlisting}

This approach is not sufficient if we want to mutate the array during iteration.
To this end, we need to take an exclusive borrow of the structure to iterate
on:
\begin{lstlisting}
val iter_mut : (&!(aff_r1,'a) -{un}> unit) -> &!('a Array.t) -> unit
\end{lstlisting}
